Bayesian regression modelling facilitates quantitative modelling of cell metabolism

### 2023-10-29 paper/main.qmd
Update paper

We conducted adaptive Hamiltonian Monte Carlo sampling for the full and missing-
-data datasets, obtaining XXX post-warmup samples after XXX of total computation
time. For the Laplace approximation comparison, we used Maud’s Laplace mode.
Full details as well as instructions for reproducing our analysis can be found
at XXX.

@@

This case study illustrates the type of kinetic model and dataset that Maud
can fit. The model we analysed has XXX reactions, XXX state variables and XXX
parameters. Posterior sampling with adaptive Hamiltonian Monte Carlo generated
XXX ESS/hour. Generalising from this result, we conclude that it is feasible
to use this method to fit models in the same order of magnitude, but not,
for example, genome-scale kinetic models. To fit larger models, faster steady
state solving methods or alternative inference algorithms will be required.
@sec-laplace addresses whether Laplace approximation is a suitable candidate.

>>>>

We conducted adaptive Hamiltonian Monte Carlo sampling for the full and missing-
-data datasets. For the full dataset we obtained 1000 post-warmup samples each
from 4 independent Markov chains after 1000 warm-up samples and "hot-starting"
with a mass metric output by a previous model run.

For the missing-data dataset XXX.

For the Laplace approximation comparison, we used Maud’s Laplace mode. Full
details as well as instructions for reproducing our analysis can be found at
XXX.

@@

This case study illustrates the type of kinetic model and dataset that Maud
can fit. The model we analysed has 10 reactions, 5 state variables and 212
parameters. Generalising from our ability to fit this model in a reasonable time
using Maud, we expect that Maud can be used to fit realistic Bayesian models
of approximately the same size, but not, for example, genome-scale kinetic
models. To fit larger models, faster steady state solving methods or alternative
inference algorithms will be required. @sec-laplace addresses whether Laplace
approximation is a suitable candidate.

### 2023-10-29 paper/main.tex
Update paper

\title[]{Bayesian regression modelling facilitates quantitative
modelling of cell metabolism}
\makeatletter
\begin{document}
\maketitle
\begin{abstract}
This paper presents Maud, a command-line application that implements
Bayesian statistical inference for kinetic models of biochemical
metabolic reaction networks. Maud takes into account quantitative
information from omics experiments and background knowledge, as well as
structural information about kinetic mechanisms, regulatory interactions
and enzyme knockouts. Below, we review the existing options in this
area, explain how Maud improves on the state of the art, describe the
intended modelling workflow and illustrate its use with an example
application.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

We conducted adaptive Hamiltonian Monte Carlo sampling for the full and
missing- -data datasets, obtaining XXX post-warmup samples after XXX of
total computation time. For the Laplace approximation comparison, we
used Maud's Laplace mode. Full details as well as instructions for
reproducing our analysis can be found at XXX.

@@

This case study illustrates the type of kinetic model and dataset that
Maud can fit. The model we analysed has XXX reactions, XXX state
variables and XXX parameters. Posterior sampling with adaptive
Hamiltonian Monte Carlo generated XXX ESS/hour. Generalising from this
result, we conclude that it is feasible to use this method to fit models
in the same order of magnitude, but not, for example, genome-scale
kinetic models. To fit larger models, faster steady state solving
methods or alternative inference algorithms will be required.
Section~\ref{sec-laplace} addresses whether Laplace approximation is a
suitable candidate.

>>>>

\title[]{Bayesian regression modelling facilitates quantitative
modelling of cell metabolism}
\makeatletter
\begin{document}
\maketitle
\begin{abstract}
This paper presents Maud, a command-line application that implements
Bayesian statistical inference for kinetic models of biochemical
metabolic reaction networks. Maud takes into account quantitative
information from omics experiments and background knowledge, as well as
structural information about kinetic mechanisms, regulatory interactions
and enzyme knockouts. Below, we review the existing options in this
area, explain how Maud improves on the state of the art, describe the
intended modelling workflow and illustrate its use with an example
application.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

We conducted adaptive Hamiltonian Monte Carlo sampling for the full and
missing- -data datasets. For the full dataset we obtained 1000
post-warmup samples each from 4 independent Markov chains after 1000
warm-up samples and ``hot-starting'' with a mass metric output by a
previous model run.

For the missing-data dataset XXX.

For the Laplace approximation comparison, we used Maud's Laplace mode.
Full details as well as instructions for reproducing our analysis can be
found at XXX.

@@

This case study illustrates the type of kinetic model and dataset that
Maud can fit. The model we analysed has 10 reactions, 5 state variables
and 212 parameters. Generalising from our ability to fit this model in a
reasonable time using Maud, we expect that Maud can be used to fit
realistic Bayesian models of approximately the same size, but not, for
example, genome-scale kinetic models. To fit larger models, faster
steady state solving methods or alternative inference algorithms will be
required. Section~\ref{sec-laplace} addresses whether Laplace
approximation is a suitable candidate.
